Different forms of uncertainty differently affect the evolution of social learning

### 2022-10-12 ms.tex
updated supplement, finished one revision

At the beginning of each generation other than the first, each social learner selects
one member of the previous generation to learn from in a payoff-biased way. 
%, if the child inherited $s_i = 1$. A child agent $c$ is a social learner if its parent $p$ was a social learner, i.e.\ $s_c \leftarrow s_p$, without mutation.
A social learner selects this ``teacher" from the previous generation by first choosing a random subset of the population and choosing among these with a probability weighted by each model's payoff (more details below).  The social learner then inherits information about the likely payoffs of each behavior from that one teacher, whereas asocial learners do not acquire this information. In this way, the social learner can potentially reduce the amount of exploration needed to both execute and learn about the optimal behavior.

@@

        $\bar\pi_{ib}$ & ``Ledger'': mean payoffs acquired via behavior $b$ by $i$ 
                       & $B$-vector of zeros \\

@@

Model dynamics proceed by first initializing the environment and agents according
to the chosen parameter settings. % for the simulation.
Then, within each generation, agents select and perform behaviors, updating their estimated payoffs (and their probabilities of choosing each behavior) along the way.
Between generations, agents reproduce, teach social learners of the next generation, and die off. 
%Child agents then begin a new round of generational dynamics, followed by reproduction. 
This process continues until the population has evolved to fixation as either all social learners or all asocial learners, or until an upper limit of 20,000 
iterations has been reached. This process is visualized in
Figure~\ref{fig:schematic}.

@@

The model is initialized with $N$ agents, all with zero observed mean payoffs, 
$\ledger = 0$, and zero counts of how many times
they have tried each behavior, $c_{ib} = 0$ (Figure~\ref{fig:schematic}A).
One of the behaviors is chosen at random to yield expected payoff $\pihigh$, with the other $B-1$ behaviors yielding
expected payoffs $\pilow$. Agents are independently randomly initialized with $s_i
\in 0,1$, so that 50\% of the initial population learn socially.

@@

Within each generation, all $N$ agents perform $L$ behaviors sequentially and
independently (Figure~\ref{fig:schematic}B).
Each agent selects one behavior on each step of its lifespan using softmax process that integrates previous knowledge. We denote the average payoffs obtained by agent $i$
for behavior $b$ as $\ledger$. We denote the number of times $i$ performed $b$
as $c_{ib}$ (``$c$'' stands for count). 
At each time step, agent $i$ performs behavior $b$ with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}}.
\end{equation}
\noindent
$\beta$ adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$). %~\citep{McElreath2005}. 
We used a default value of $\beta = 10$, but also varied this value in our sensitivity analyses.

Agent $i$ using behavior $b$ receives a payoff of 1 with probability $\pi_b$. 
%Agents probabilistically recieve a payoff of 0 or 1 depending on whether the bandit paid off. Behavior payoffs are drawn from a Bernoulli distribution with mean $\pi_b$, so $\pi_b$ is also the mean payoff from behavior $b$.  For concreteness we denote this probabilistic payoff $\bandit$. 
After performing behavior $b$, agent $i$ updates its
corresponding \emph{behavior count} for $b$ by 1, $c_{ib}' \leftarrow c_{ib} + 1$.
Then the observed average payoffs known by agent $i$ for behavior $b$ are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger' \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}'},
\end{equation}
\noindent
where $\bandit$ is the actual payoff received by agent $i$ for behavior $b$ on the
given (implicit) time step.

\paragraph{Intergenerational dynamics} Between generations, agents first reproduce
via asexual haploid reproduction, which determines the transmission of the social
learning trait. Social learners then learn from the previous generation, after which
all agents from the previous generation die off (Figure~\ref{fig:schematic}C). More
specifically, this all happens as follows.  $N$ reproducers are selected with
replacement over $N$ independent draws, biased by performance: \begin{equation}
\Pr(\text{Agent $i$ is chosen to reproduce in one draw}) =
\frac{\pi_i}{\sum_{i=1}^N \pi_i}.  \end{equation} \noindent A child inherits its
parent's social learning trait $s_i$ without mutation.  A social learner child with
$s_i = 1$ learns from a teacher from its parent's generation, including possibly
their parent (i.e., both vertical and oblique learning are possible).  A child
chooses its teacher by first randomly selecting $N_T$ prospective teachers from the
population, then selecting the one in this set with the greatest payoff net, with ties broken randomly. By first subsetting $N_T$
prospective teachers we represent the fact that access to the entire population is
not generally guaranteed, though this type of algorithm yields qualitatively
similar results to ones that do not first produce subsamples
\citep{smaldino2019open} and we show in the supplement 
that our results are robust to different $N_T$
(Figure~\ref{fig:nteachersSensitivity}).  Social learner $i$ adopts their chosen 
teacher $j$'s observed average payoffs (average indicated by the overbar) 
for each behavior, $b$, and $c_{ib}$ is set to 1 if teacher $j$ had at least one
observation of behavior $b$,  
and $c_{ib}$ is set to 0 otherwise. This way, expected payoffs remain within $[0,
1]$, but social learners are flexible to adjust to environmental change. 
Setting $c_{ib}$ to 1 reflects the fact that the social learning is 
treating the teacher's information as a single observation with the associated
uncertainty. Asocial learners ($s_i = 0$) are initialized with $c_{ib} = 0$ and
$\ledger = 0$ for each behavior, $b$.
(Figure~\ref{fig:schematic}C). The newly initialized population
then repeats the within-generation dynamics.

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values (Table~\ref{tab:modelParameters}) and observed how frequently social learning evolved across
1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
We observed three outcome measures, with $\meansl$ as the primary measure: the average value of $s_i$ over all agents and trials. To support our conclusions
about $\meansl$, we also measured $\meanpi / L$, the average net  payoffs at the end of the simulation across agents and trials, normalized by lifespan; 
and $\meanG$, the mean
number of generations to fixation (see Table~\ref{tab:outcomeVariables}). We analyze outcomes by plotting $\meansl$ on the y-axis and environmental variability is on x-axis since we
theoretically expect that $\meansl$ will decrease monotonically from 1 to 0 as $u$
increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. These analyses are presented in 
supplemental material. First, we varied the population size, $N$, to explore whether smaller $N$ would show more drift due to finite population size effects. %, but  still generally support our main conclusions. 
Next, we varied the number of prospective teachers, $N_T \in \{2, 10, 20\}$, to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change whether social learning is optimal, but it may induce more drift since the benefit of social learning may
be more difficult to detect with fewer prospective teachers. Finally, we varied the
softmax greediness parameter $\beta \in \{1, 100\}$ to supplement the main results
that used $\beta = 10$. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
by undermining individual learning. Nonetheless, our results
are valid over a substantial range of $\beta$.

@@

Our model was implemented in the Julia programming language~\citep{Bezanson2017} 
using the Agents.jl agent-based modeling library~\citep{Datseris2022} and run
on the Sherlock supercomputing cluster at Stanford University. Model code and
data are publicly available on GitHub at \url{https://github.com/mt-digital/UncMod}.

@@

Our main results are illustrated by Figure~\ref{fig:mainResults}, which shows the proportion of simulation runs that fixated to 100\% social learners as a function of each of our four uncertainty measures.  First, we find that reliance on social learning monotonically decreases as temporal environmental variability, $u$, increases. When the environment is more likely to change between generations, information learned from the previous generation is less likely to be of value, decreasing selection for social learning. Past some threshold, social learning is not favored at all. However, the exact nature of the relationship between social learning and temporal environmental variability is moderated by our other three uncertainty parameters.

@@

Of the four types of uncertainty studied in this paper, payoff ambiguity has the most
complex effect on the evolution of social learning. Recall that we operationalize
payoff ambiguity with $\pilow$, the expected payoff for non-optimal behaviors
(equivalently the probability that those behaviors yielded a payoff of 1). This was relative to $\pihigh = 0.9$, the expected payoff for the optimal behavior. When payoff ambiguity was low, non-optimal behaviors rarely paid off, so that exploration yielded reliable information and selection on learning strategies was  strong. When payoff ambiguity is very high ($\pilow = 0.8$), two things happen. First, agents are more likely to err by ascribing high value to non-optimal behaviors, since it is more difficult for them to discern the difference between optimal and sub optimal behaviors. Second, and perhaps more importantly, natural selection on strategies that $do$ reliably select the optimal behavior is relatively weak. This means that the effect of drift is relatively stronger, weakening selection in favor of social learning under small values of $u$ and weakening selection against social learning under high values of $u$ (Figure~\ref{fig:steps}). This effect was especially strong under other types of uncertainty, i.e., larger selection set sizes and especially shorter effective lifespans---particularly when $L=1$ where no individual learning occurred.

The effect of drift when effective lifespans are short and payoffs are very ambiguous is powerful and complex. We explain how the frequency of social learning evolves in such circumstances in the following section.
%selection on social learning responds to these effects in more detail in the following section. 
%\cm{I rephrased as I didn't think selection was RESPONDING to drift. Original framing sounded weird since in essence we have weak selection which is why drift is relatively high. }

@@

$\meanG$ peaks at the values of $u$ where social learning becomes suppressed. $B$, and $\pilow$ have the
clearest effects on the location of greatest drift over $u$. $L$ has some effect on the location of peak
drift over $u$, but $L$ also has a large effect on the overall amount of drift: $L=1$ tends to have the most
drift with sufficiently low payoff ambiguity ($\pilow < 0.8$). The effect of lifespan on drift is
non-monotonic, since we observe that long lifespans can also result in relatively greater drift
compared to lifespans other than $L=1$ (e.g., in Figure~\ref{fig:steps}, top left and right).
When $L=1$ selection pressure may be relatively weak since agents have no opportunity
for individual learning, and social learning is the only learning channel. 
When $L=8$ ($B=2,4$) or when $L=20$ ($B=10$), individual learning has many 
opportunities for finding the optimal behavior, which means social learning
provides a relatively weaker constraint on net payoffs over agent lifespans.
This also results in weaker selection pressure since there is relatively little
difference between social and asocial constraints when $L$ is large.

@@

Figure~\ref{fig:payoffs} shows the average net payoffs normalized by lifespan
for our simulations and two reference conditions where agents are initialized
as homogenous populations fixated either as all social or all asocial learners.
Often the simulated payoffs follow the payoffs from the better-performing
homogenous group, but there are several important deviations to understand.
First, we observed frequency dependence of social learning, where 
agents from our main simulations who evolved to be social learners 
outperform agents in the homogenous social learner populations. This is most
pronounced when $L=1$, across all other uncertainty parameters, and even more
pronounced when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}, middle right).
In these cases, the presence of asocial learners before social learning fixates
enabled social learners in our simulations to outperform all-social learner populations.
Frequency dependence of social learning payoffs even pushed social learning to evolve even
when the expected homogenous social learning payoff was less than the expected homogenous
asocial learning payoff (Figure~\ref{fig:payoffs}, middle center $L=1$).
We also observed cases where long-lived populations ($L=4$ and $L=8$) in simple environments
($B=2$) evolved to be individual learners, even though all-social learner populations
outperformed all-individual learner populations (Figure~\ref{fig:payoffs}, upper left).

We explain these deviations from expected homogenous payoffs 
by inspecting time series of individual model runs where we plot the geometric moving averages (with time window 3) of the payoffs for the whole population, and for the series of payoffs broken out by asocial/social learner sub-populations, and compare these series to the expected homogenous social and asocial payoffs, the time series of social learner prevalence, and to the timing of environmental changes within the simulation (Figure~\ref{fig:payoffTimeseries}). 
When lifespan was minimal ($L=1$) and environmental variability
was moderately high ($u\approx0.5$), inspection of payoffs a period of environmental
stability led populations to fixate to social learning, despite individual learning
being the superior long-term strategy (Figure~\ref{fig:payoffs}, center top and middle).
This behavior was again confirmed by examining time series of simulations
initialized with the commensurate uncertainty parameters, which show that
social learning fixates during periods of protracted environmental stability
(Figure~\ref{fig:payoffTimeseries}a).
When payoff ambiguity was large, decision set size was small $B=2$, and lifespan was long ($L=4$ and $L=8$),
asocial learning fixates as a hedge against periods of protracted environmental
instability, which causes social learning to frequently provide
outdated information, even though expected homogenous social payoffs are greater than
homogenous asocial payoffs (Figure~\ref{fig:payoffTimeseries}b,c).

@@

Our modeling results suggest, on the one hand, that society might be better to ignore social information from previous generations since our natural world is rapidly changing~\citep{IPCC2022}; on the other hand, however, humans have increasingly more behavioral options (decision set size) with high levels of uncertainty about which behaviors are most beneficial (payoff ambiguity), and in this case our model predicts a greater prevalence of social learning.
In general, detailed theoretical models like this are likely to be critical for understanding how humans might adapt to an uncertain and rapidly changing world.
Such understanding could suggest plausibly beneficial interventions to mitigate existential threats~\citep{Moya2020,Jones2021}, especially in
the most vulnerable communities~\citep{McNamara2020}, and to capitalize on new behavioral opportunities such as transitioning to clean energy use and
production~\citep{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}. However, to better identify potential pathways to effective adaptation it may be necessary to adjust some of our modeling assumptions to include both the cumulative nature of cultural evolution and group structure of interpersonal exchange, both of which have serious consequences for understanding the sustainability of novel adaptations~\citep{Derex2020,Centola2018,Smaldino2021a}.
\cm{should we draw out some implications? I find this difficult, but e.g. stressing environmental change may discourage learning from older gen? having more tech options available for dealing with climate change may encourage social learning? Frankly so many of these problems seem to come back to encouraging people to socially learn from the RIGHT people, not just socially or asocially learning.} \mt{I gave this a try, with some restructuring to put the note about social learning as a scaffold for individual learning in the above paragraph.}

@@

We could have made alternative, empirically-valid, choices for a range of other structural features, the future exploration of which will be important to the development of a more nuanced theory of social learning.  % that may not always hold in more specific situations, though they are empirically justified to be general across taxa. 
For example, we only considered success-biased
learning, although people may use conformity or other context biases to socially learn ~\citep{BoydRicherson1985,Muthukrishna2016a,Smaldino2018b}.  This modelling choice maximizes the potential benefit to social learners suggesting other mechanisms, like conformity, may lead to more drift overall. We also assumed that the number of behaviors and
their payoffs were constant within a given simulation, but this fails to account for evolutionary feedback which creates new behavioral opportunities as time progresses, e.g.\ via niche construction~\citep{Smaldino2012a,Heras-Escribano2020} or cumulative cultural evolution
%supported by dynamic network structure
~\citep{Smolla2019,Derex2020}.  Group structure
and processes such as homophily and in-group bias/out-group aversion could inhibit the
evolution of social learning since successful out-group teachers could be cut off from
in-group learners~\citep{Golub2012}. We assumed agents choose behaviors via softmax
search, but even though this algorithm is more sophisticated than most used in social learning models, it is algorithm compared to real human learning mechanisms ~\citep{Schulz2020a,Wu2022}. Instantiating even more powerful asocial learning could further
support the evolution of social learning. These considerations reveal new opportunities to enhance theoretical nuance by adjusting or extending the present model for other contexts.

@@

Social learning is a critical component of problem solving among social creatures. Our work identified and clarified some important forms of uncertainty and catalogued their effects on 
model populations' collective decisions to learn socially or not when collectively problem solving. Our work, then, advances both the theory of social learning and the theory of collective problem solving more generally, and could eventually be of practical use in designing sustainable adaptations to the immense threats and uncertainties faced by humans today.
% This work most directly informs our current understanding of the evolution of social learning, but also has broader significance for understanding problem solving under uncertainty. 
\ps{What is this significance? Can't just leave that dangling. I also cut the bit about what comes next. No need for spoilers.} \mt{Gave this a try, but I think Jamie should have a crack at this.}
%Our model and its software implementation were designed to be modular and extensible.  Indeed, we plan to extend this model to further develop a more detailed theoretical understanding of the evolution of social learning; we hope others will, too.

@@

\begin{figure}
  \centering
  % \addtocounter{figure}{-1}
  \caption{
	Sensitivity analysis of the main results for different population
	sizes, $N=50,200,1000$. Recall $N=100$ was used to generate main 
	text results.
  }
  \label{fig:populationSensitivity}
  \begin{subfigure}{\textwidth}
	\caption{$N=50$}
	\includegraphics[width=\textwidth]{Figures/supplement/nagents=50/mainResultsPlots.pdf}
  \end{subfigure}
\end{figure}

\begin{figure}
  \ContinuedFloat
	\begin{subfigure}{\textwidth}
	  \caption{$N=100$}
	  \includegraphics[width=\textwidth]{Figures/supplement/nagents=100/mainResultsPlots.pdf}
	\end{subfigure}
\end{figure}

>>>>

At the beginning of each generation other than the first, each social learner selects
one member of the previous generation to learn from in a payoff-biased way. 
%, if the child inherited $s_i = 1$. A child agent $c$ is a social learner if its parent $p$ was a social learner, i.e.\ $s_c \leftarrow s_p$, without mutation.
A social learner selects this ``teacher" from the previous generation by first
choosing the payoff maximizer among these (more details below).  The social learner then inherits information about the likely payoffs of each behavior from that one teacher, whereas asocial learners do not acquire this information. In this way, the social learner can potentially reduce the amount of exploration needed to both execute and learn about the optimal behavior.

@@

        $\bar\pi_{ib}$ & Mean payoffs acquired by agent $i$ from behavior $b$
                       & $B$-vector of zeros \\

@@

Model dynamics proceed by first initializing the environment and agents according
to the chosen parameter settings. % for the simulation.
Then, within each generation, agents select and perform behaviors, updating their estimated payoffs (and their probabilities of choosing each behavior) along the way.
Between generations, agents reproduce, teach social learners of the next generation, and die off. 
%Child agents then begin a new round of generational dynamics, followed by reproduction. 
This process continues until the population has evolved to fixation as either all social learners or all asocial learners. This process is visualized in
Figure~\ref{fig:schematic}.

@@

The model is initialized with $N$ agents. Each agent $i$ tracks 
observed mean payoffs for each behavior, $b$, which is denoted $\ledger$. Each
agent $i$ also counts how many times they have tried each behavior $b$, 
denoted $c_{ib}$. At model initialization, $\ledger = 0.0$ and $c_{ib} = 0$ for
all $i$ and $b$ (Figure~\ref{fig:schematic}A).
One of the behaviors is chosen at random to yield expected payoff $\pihigh$, with the other $B-1$ behaviors yielding
expected payoffs $\pilow$. Agents are independently randomly initialized with $s_i
\in 0,1$, so that $\sim$50\% of the initial population learn socially.

@@

Within each generation, all $N$ agents perform $L$ behaviors sequentially and
independently (Figure~\ref{fig:schematic}B).
% Each agent $i$ selects one behavior on each step of its lifespan using softmax applied
% to the vector of mean payoffs the agent has observed for each behavior
% ($\ledger$), integrating previous knowledge. 
% We denote the average payoffs obtained by agent $i$
% for behavior $b$ as $\ledger$. We denote the number of times $i$ performed $b$
% as $c_{ib}$ (``$c$'' stands for count). 
At each time step, agent $i$ performs behavior $b$ with softmax-weighted probability
\begin{equation}
  \Pr(\text{Agent $i$ performs behavior $b$}) = 
    \frac{e^{\beta \ledger}}{\sum_{b=1}^B e^{\beta \ledger}}.
\end{equation}
\noindent
$\beta$ adjusts how frequently agents perform 
behaviors with high expected payoffs (larger $\beta$) versus how frequently
agents explore alternative behaviors (smaller $\beta$). %~\citep{McElreath2005}. 
We used a default value of $\beta = 10$, but also varied this value in our
sensitivity analyses~(Figure~\ref{fig:softmaxSensitivity}).

Agent $i$ using behavior $b$ receives a payoff of 1 with probability $\pihigh$
if $b$ is the optimal behavior, or with probability $\pilow$ otherwise. 
%Agents probabilistically recieve a payoff of 0 or 1 depending on whether the bandit paid off. Behavior payoffs are drawn from a Bernoulli distribution with mean $\pi_b$, so $\pi_b$ is also the mean payoff from behavior $b$.  For concreteness we denote this probabilistic payoff $\bandit$. 
After performing behavior $b$, agent $i$ updates its
corresponding \emph{behavior count} for $b$ by 1, $c_{ib}' \leftarrow c_{ib} + 1$.
Then the observed average payoffs known by agent $i$ for behavior $b$ are updated with
exponentially-weighted averaging
\begin{equation}
  \ledger' \leftarrow \ledger + \frac{\bandit - \ledger}{c_{ib}'},
\end{equation}
\noindent
where $\bandit$ is the actual payoff received by agent $i$ for behavior $b$ on the
given (implicit) time step.

\paragraph{Intergenerational dynamics} Between generations, agents first reproduce
via asexual haploid reproduction, which determines the transmission of the social
learning trait. Social learners then learn from the previous generation, after which
all agents from the previous generation die off (Figure~\ref{fig:schematic}C). More
specifically, this all happens as follows.  $N$ reproducers are selected with
replacement over $N$ independent draws, biased by performance: 
\begin{equation}
  \Pr(\text{Agent $i$ is chosen to reproduce in one draw}) =
\frac{\pi_i}{\sum_{i=1}^N \pi_i}.  
\end{equation} 
\noindent 
A child inherits its
parent's social learning trait $s_i$ without mutation.  A social learner child with
$s_i = 1$ learns from a teacher from its parent's generation, including possibly
their parent (i.e., both vertical and oblique learning are possible).  A child
chooses its teacher by first randomly selecting $N_T$ prospective teachers from the
population, then selecting the one in this set with the greatest payoff net, with ties broken randomly. By first subsetting $N_T$
prospective teachers we represent the fact that access to the entire population is
not generally guaranteed, though this type of algorithm yields qualitatively
similar results to ones that do not first produce subsamples
\citep{smaldino2019open} and we show in the supplement 
that our results are robust to different $N_T$
(Figure~\ref{fig:nteachersSensitivity}).  Social learner $i$ adopts their chosen 
teacher $j$'s observed average payoffs for each behavior, $b$, i.e. 
$\ledger \leftarrow \bar\pi_{jb}$; 
and agent $i$'s count of behavioral observations, $c_{ib}$, is set to 1 if teacher $j$ had at least one observation of behavior $b$,  
and $c_{ib}$ is set to 0 otherwise. This way, expected payoffs remain within $[0,
1]$, but social learners are flexible to adjust to environmental change. 
Setting $c_{ib}$ to 1 reflects the fact that the social learning is 
treating the teacher's information as a single observation with the associated
uncertainty. Asocial learners ($s_i = 0$) are initialized with $c_{ib} = 0$ and
$\ledger = 0$ for each behavior, $b$.
(Figure~\ref{fig:schematic}C). The newly initialized population
then repeats the within-generation dynamics.

@@

To analyze the effect of the four principal uncertainty factors, we systematically
varied uncertainty values (Table~\ref{tab:modelParameters}) and observed how frequently social learning evolved across
1000 trial simulations for each combination of parameters tested. %We varied $u \in \{0.0, 0.1, \ldots, 1.0\}$; $\pilow \in \{0.1, 0.45, 0.8\}$; $B \in \{2, 4, 10\}$; and $L \in \{1,2,4,8\}$ for $B=2$ and $L \in \{1,B/2,B,2B\}$ for $B=4,10$.  
We observed three outcome measures, with $\meansl$ as the primary measure: the average value of $s_i$ over all agents and trials. To support our conclusions
about $\meansl$, we also measured $\meanpi / L$, the average net  payoffs at the end of the simulation across agents and trials, normalized by lifespan; 
and $\meanG$, the mean
number of generations to fixation (see Table~\ref{tab:outcomeVariables}). We analyze
outcomes by plotting $\meansl$ on the y-axis and environmental variability on
x-axis, since we
theoretically expect that $\meansl$ will always decrease monotonically from 1 to 0 as $u$
increases. %Note that the hypothesis-testing concept of significance is meaningless here because we could make any small outcome difference ``significant'' by running more simulation trials.  We instead study the patterns of outcome variables over systematically varied uncertainty factor values.

@@

We performed sensitivity analyses to ensure that our main analysis is reasonably
robust to a range of auxiliary parameters. These analyses are presented in 
supplemental material. First, we varied the population size, $N$, to confirm that
smaller $N$ would make similar predictions, but with more drift due to finite population size effects. %, but  still generally support our main conclusions. 
Next, we tested model robustness to the number of prospective teachers, $N_T \in \{2, 10, 20\}$, to supplement the main results that used $N_T = 5$. The number of prospective
teachers should not change whether social learning is optimal, but it may induce more drift since the benefit of social learning may
be more difficult to detect with fewer prospective teachers. Finally, we tested
model robustness to varying the
softmax greediness parameter $\beta \in \{1, 100\}$, to supplement the main results
that used $\beta = 10$. We do not
expect agents with different $\beta$ to perform equally well 
individually, so certain $\beta$ values may themselves suppress social learning
by improving or undermining individual learning.

@@

Our model was implemented in the Julia programming language~\citep{Bezanson2017} 
using the Agents.jl agent-based modeling library~\citep{Datseris2022} and run
on the Sherlock supercomputing cluster at Stanford University. Model code is
publicly available on GitHub at \url{https://github.com/mt-digital/UncMod},
and output data used to generate the figures presented here are available on OSF.io
\mt{TODO: \url{https://make.OSF.repo}}.

@@

Our main results are illustrated by Figure~\ref{fig:mainResults}, which shows the
proportion of simulation runs that fixated to 100\% social learners as a function of
each of our four uncertainty measures.  First, we found that reliance on social
learning monotonically decreases as temporal environmental variability, $u$,
increases. When the environment is more likely to change between generations,
information learned from the previous generation is less likely to be of value,
decreasing selection for social learning. Past some threshold, social learning is
not favored at all. However, the exact nature of the relationship between social
learning and temporal environmental variability was moderated by the other three uncertainty parameters.

@@

Of the four types of uncertainty studied in this paper, payoff ambiguity had the most
complex effect on the evolution of social learning. Recall that we operationalized
payoff ambiguity with $\pilow$, the expected payoff for non-optimal behaviors
(equivalently the probability that those behaviors yielded a payoff of 1). This was
relative to $\pihigh = 0.9$, the expected payoff for the optimal behavior. When
payoff ambiguity was low, non-optimal behaviors rarely paid off, so that exploration
yielded reliable information and selection on learning strategies was  strong. When
payoff ambiguity is very high ($\pilow = 0.8$), two things happen. First, agents are
more likely to err by ascribing high value to non-optimal behaviors, since it is
more difficult for them to discern the difference between optimal and sub optimal
behaviors. Second, and perhaps more importantly, natural selection on strategies
that $do$ reliably select the optimal behavior is relatively weak. This means that
the effect of drift is relatively stronger, weakening selection in favor of social
learning under small values of $u$ and weakening selection against social learning
under high values of $u$ (Figure~\ref{fig:mainResults}, bottom row). 
This effect was especially strong under other types of uncertainty, i.e., larger selection set sizes and especially shorter effective lifespans---particularly when $L=1$ where no individual learning occurred.

The effect of drift when the benefit of social learning is ambiguous is powerful and complex. We explain how the frequency of social learning evolves in such circumstances in the following section.
%selection on social learning responds to these effects in more detail in the following section. 
%\cm{I rephrased as I didn't think selection was RESPONDING to drift. Original framing sounded weird since in essence we have weak selection which is why drift is relatively high. }

@@

$\meanG$ peaks at the values of $u$ where social learning becomes suppressed. 
$B$ and $\pilow$ have the
clearest effects on the location of greatest drift over $u$. $L$ has some effect on the location of peak
drift over $u$, but $L$ also has a large effect on the overall amount of drift: $L=1$ tends to have the most
drift with sufficiently low payoff ambiguity ($\pilow < 0.8$). The effect of lifespan on drift is
non-monotonic, since we observe that long lifespans can also result in relatively greater drift
compared to lifespans other than $L=1$ (e.g., in Figure~\ref{fig:steps}, top left and right).
When $L=1$ selection pressure may be relatively weak since agents have no opportunity
for individual learning, and social learning is the only learning channel. 
When $L=8$ ($B=2,4$) or when $L=20$ ($B=10$), individual learning has many 
opportunities for finding the optimal behavior, which means social learning
provides a relatively weaker constraint on net payoffs over agent lifespans.
This also results in weaker selection pressure since there is relatively little
difference between social and asocial constraints when $L$ is large
(Figure~\ref{fig:steps}, top two rows).

@@

Figure~\ref{fig:payoffs} shows the average net payoffs normalized by lifespan
for our simulations compared with the two reference homogenous populations
initialized either as all social or all asocial learners.
Often the simulated payoffs follow the payoffs from the better-performing
homogenous group, but there are several important deviations to understand.
First, we observed frequency dependence of social learning, where 
agents from our main simulations who evolved to be social learners 
outperform agents in the homogenous social learner populations. This is most
pronounced when $L=1$, across all other uncertainty parameters, and even more
pronounced when $\pilow=0.45$ and $B=10$ (Figure~\ref{fig:payoffs}, middle right).
In these cases, the presence of asocial learners before social learning fixated
enabled social learners in our simulations to outperform all-social learner populations.
Frequency dependence of social learning payoffs even pushed social learning to evolve even
when the expected homogenous social learning payoff was less than the expected homogenous
asocial learning payoff (Figure~\ref{fig:payoffs}, middle center $L=1$).
We also observed cases where long-lived populations ($L=4$ and $L=8$) in simple environments
($B=2$) evolved to be individual learners, even though all-social learner populations
outperformed all-individual learner populations (Figure~\ref{fig:payoffs}, upper left).

We explain these deviations from expected homogenous payoffs 
by inspecting time series of individual model runs where we plot the geometric
moving averages (with time window 3) of the payoffs for the whole population, and
for the series of payoffs broken out by asocial/social learner sub-populations; we
compare these mean payoff series to the expected homogenous social and asocial payoffs, the time series of social learner prevalence, and to the timing of environmental changes within the simulation (Figure~\ref{fig:payoffTimeseries}). 
When lifespan was minimal ($L=1$) and environmental variability
was moderate ($u\approx0.5$), the payoff time series show a period of environmental
stability led populations to fixate to social learning, despite individual learning
being the superior long-term strategy (Figure~\ref{fig:payoffTimeseries}a).
When payoff ambiguity and decision set size was small ($\pilow=0.1$ and $B=2$), 
and lifespan was long ($L=4$ and $L=8$),
asocial learning fixated as a hedge against periods of protracted environmental
instability, which caused social learning to frequently provide
outdated information, even though expected homogenous social payoffs are greater than
homogenous asocial payoffs (Figure~\ref{fig:payoffTimeseries}b,c).

@@

Our modeling results suggest, on the one hand, that society might be better to
ignore social information from previous generations since our natural world is
rapidly changing~\citep{IPCC2022}; on the other hand, humans have increasingly more
behavioral options (selection set size) with high levels of uncertainty about which behaviors are most beneficial (payoff ambiguity), and in this case our model predicts a greater prevalence of social learning.
In general, detailed theoretical models like this are likely to be critical for understanding how humans might adapt to an uncertain and rapidly changing world.
Such understanding could suggest plausibly beneficial interventions to mitigate existential threats~\citep{Moya2020,Jones2021}, especially in
the most vulnerable communities~\citep{McNamara2020}, and to capitalize on new behavioral opportunities such as transitioning to clean energy use and
production~\citep{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}. However, to better identify
potential pathways to effective adaptation it may be necessary to adjust some of our modeling assumptions to
include both the cumulative nature of cultural evolution and group structure of interpersonal exchange, both
of which have been shown to affect the sustainability of novel adaptations~\citep{Derex2020,Centola2018,Smaldino2021a}.
\cm{should we draw out some implications? I find this difficult, but e.g. stressing environmental change may discourage learning from older gen? having more tech options available for dealing with climate change may encourage social learning? Frankly so many of these problems seem to come back to encouraging people to socially learn from the RIGHT people, not just socially or asocially learning.} \mt{I gave this a try, with some restructuring to put the note about social learning as a scaffold for individual learning in the above paragraph.}

@@

We could have made alternative, empirically-valid, choices for a range of other structural features, the future exploration of which will be important to the development of a more nuanced theory of social learning.  % that may not always hold in more specific situations, though they are empirically justified to be general across taxa. 
For example, we only considered success-biased
learning, although people may use conformity or other context biases to socially
learn~\citep{BoydRicherson1985,Muthukrishna2016a,Smaldino2018b}.  This modelling
choice maximizes the potential benefit to social learners, which suggests that
other mechanisms, like conformity, may lead to more drift overall. We also assumed that the number of behaviors and
their payoffs were constant within a given simulation, but this fails to account for evolutionary feedback which creates new behavioral opportunities as time progresses, e.g.\ via niche construction~\citep{Smaldino2012a,Heras-Escribano2020} or cumulative cultural evolution
%supported by dynamic network structure
~\citep{Smolla2019,Derex2020}.  Group structure
and processes such as homophily and other group-level biases could inhibit the
evolution of social learning since successful out-group teachers could be cut off from
in-group learners~\citep{Golub2012}. We assumed agents choose behaviors via softmax
search, but even though this algorithm is more sophisticated than most used in social learning models, it is algorithm compared to real human learning mechanisms ~\citep{Schulz2020a,Wu2022}. Instantiating even more powerful asocial learning could further
support the evolution of social learning. These considerations reveal new opportunities to enhance theoretical nuance by adjusting or extending the present model for other contexts.

@@

Social learning is a critical component of problem solving among social creatures.
Our work identified and systematized some important forms of uncertainty and catalogued their effects on 
the evolution of social learning. Our work, then, both advances the theory of social
learning and could eventually be of practical use in designing sustainable
adaptations to an uncertain future.
% This work most directly informs our current understanding of the evolution of social learning, but also has broader significance for understanding problem solving under uncertainty. 
\ps{What is this significance? Can't just leave that dangling. I also cut the bit about what comes next. No need for spoilers.} \mt{Gave this a try, but I think Jamie should have a crack at this.}
%Our model and its software implementation were designed to be modular and extensible.  Indeed, we plan to extend this model to further develop a more detailed theoretical understanding of the evolution of social learning; we hope others will, too.

@@

\begin{figure}
  \centering
  % \addtocounter{figure}{-1}
  \caption{
	Sensitivity analysis of the main results for two population
	sizes, $N=50,200$. Recall $N=1000$ was used to generate main 
	text results.
  }
  \label{fig:populationSensitivity}
  \begin{subfigure}{\textwidth}
	\caption{$N=50$}
	\includegraphics[width=\textwidth]{Figures/supplement/numagents=50/mainResultsPlots.pdf}
  \end{subfigure}
\end{figure}

\begin{figure}
  \ContinuedFloat
	\begin{subfigure}{\textwidth}
	  \caption{$N=200$}
	  \includegraphics[width=\textwidth]{Figures/supplement/numagents=200/mainResultsPlots.pdf}
	\end{subfigure}
\end{figure}
